---
title: 'Module 6 exercises'
output: html_notebook
---

```{r}
library(ggplot2) #load ggplot
library(ggthemes) #load themes package for ggplot
theme_set(theme_tufte()) # a theme I like.

#set some options that I like
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```

---------------
# Exercise 6A #
---------------

## Examining multicollinearity

Reproduce the monet VIF analysis from lecture. Compare the VIF's from the full model ('WIDTH + HEGHT + SIZE') against a model with `WIDTH + SIZE`. Did removing `HEIGHT` improve the multicollinearity? Does this match what you would expect?

```{r}
#code here
library(car)
monet.dat <- read.csv(file="../Data/monet.csv")

full.mod <- lm(log(PRICE) ~ HEIGHT + WIDTH + SIZE, data=monet.dat)
red.mod <- lm(log(PRICE) ~ WIDTH + SIZE, data=monet.dat)
vif(red.mod)
```
# Removing height improved multicollinearity but not as much as removing SIZE.

---------------
# Exercise 6B #
---------------

## Adjusted $R^2$

Add an additional model to the analyses from class using a model with just the `Danger` predictor variable. Does this model perform better or worse than the models from class when using adjusted $R^2$ to assess model fit?

```{r}
sleep.dat <- read.csv(file="../Data/MammalSleep.csv")
sleep.dat <- sleep.dat[complete.cases(sleep.dat), ] #remove NA's

# code here
mod4 <- lm(TotalSleep ~ Danger, data=sleep.dat)
summary(mod4)
```

### This model performs worse than the models explored in class in terms of the adjusted $R^2$.

----------------
# Excercise 6C #
----------------

## LOOCV

Apply LOOCV to the Danger model from 6B. How does it perform relative to the models in class?

```{r}
library(caret)
#code here
sleep.control <- trainControl(method = "LOOCV") #tell the function we want to do leave one out cross validation
mod4.xv <- train(TotalSleep ~ Danger, data=sleep.dat, method = "lm", na.action=na.omit, trControl=sleep.control)
mod4.xv$results
```
### This model performs worse than the ones explored in lecture.

----------------
# Excercise 6D #
----------------

## Model selection with AIC

Calculate the AICc value for the Danger model  using the `MuMIn` package from problem 6C. Does this agree with your assessments using the adjusted $R^2$ and LOOCV?

```{r}
library(MuMIn)
#code here
model.sel(mod4)
```

### This model performs worse than all other models explored in lecture in terms of AICc.
