---
title: 'Module 7 exercises'
output: html_notebook
---

```{r}
library(ggplot2) #load ggplot
library(ggthemes) #load themes package for ggplot
theme_set(theme_tufte()) # a theme I like.

#set some options that I like
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```

---------------
# Exercise 7A #
---------------

## Fit a count model

Below is the code I used to generate the simualated pheasant data from class. Run it and then fit a poisson glm using the formuma `y~x`. Do you get a similar result from lecture? Run the chunk a couple of times to see how much variability there is in your regression estimates.

```{r}
x<-seq(0,100,1) 
mus<-exp(0.01+0.03*x) # mean at each x
y<-rpois(length(x), lambda= mus) # 1 observation from a poisson per x value

#code here
pois.mod <- glm(y ~ x, family=poisson)
pois.mod
```
### Everytime I run the chunk I get a slightly different intercept and slope term.


---------------
# Exercise 7B #
---------------

## Negative binomial regression

Repeat the analysis in 7A but use a negative binomial model. Use AICc to determine which fit is better. Do the parameter estimates or confidence intervals (from `confint`) differ much between the results in 7A and 7B?

```{r}
library(MASS)
library(MuMIn)
#code here
nb.mod <- glm.nb(y ~ x)
model.sel(list(nb.mod, pois.mod))
```
### Based on AICc the Poisson model is better.

---------------
# Exercise 7C #
---------------

## Modeling categorical data

Let's look at the silversword data from the lecture. 

1. Load the dataset `Silversword.csv`. Bonus points if you can figure out how to plot the survival data.
2. Build a survival model using the predictor `Precip` (instead of the `Elev` predictor we used in class). Get the AICc value for this model
3. Based on this model is survival related to `Precip`? If so, does it increase or decrease with `Precip`?

```{r}
#code here
silver.dat <- read.csv(file="../Data/Silversword.csv")

silver.mod <- glm(cbind(Survive, Die) ~ Precip, data=silver.dat, family=binomial)
summary(silver.mod)
model.sel(silver.mod)
```

### Based on this model survival declines with precipitation. 

---------------
# Exercise 7D #
---------------

## Fit an overdispersed categorical model

1. Use the `VGAM` package to fit the model you used above to fit the same mean function as in 7C but using a betabinomial model. 
2. Compare the AICc values between the two models and compare their parameter estimates , are there differences in the estimates? Note that you can use AICc(model.name) to extract the AICc value for the betabinomial model because the MuMIn library doesn't work on models fit using the VGAM package.

```{r}
library(VGAM)
#code here
silver.bb <- vglm(cbind(Survive, Die) ~ Precip, data=silver.dat, family=betabinomial)
summary(silver.bb)
AICc(silver.bb)
```
### The beta-binomial model has much more support than the binomial model based on the AICc values. In this updated model, the effect of precipitaiton cannot be distinguished from 0 (p-value=0.392)


