---
title: 'Module 3 group work'
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(ggplot2) #load ggplot
library(ggthemes) #load themes package for ggplot
theme_set(theme_tufte()) # a theme I like.

#set some options that I like
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```

---------------
# Exercise 3A #
---------------

## Using the sample function

Sample with replacement from -1, 1 as in lecture. First try a sample size of 5 and get the mean and standard deviation. Run this code a couple of times. Do these statistics change? Now do the same thing with a sample size of 1000. Do these statistics change?
```{r}
#code here
```


---------------
# Exercise 3B #
---------------

## Your first for loop

Loops allow you to repeat a chunk of code a bunch of times.

syntax: `for(index in set) {do this}`
Here is a couple of examples for you to explore:
```{r}
#For loop syntax: for (var in seq) {expression}
for (i in 1:10) {
  print(i)
} #basic for loop

```

```{r}
for (i in c(1,3,17)) {
  print(i)
} #sequence need not be in order
```

```{r}
for (char in c("A", "Z", "2")) {
  print(char)
} #sequence need not be numbers
```

Below is the loop illustrated  in class.
1. Change the index `i` to be called `x`. 
2. Change the sequence from 1 - 5 to 101-105.

```{r}
#update this code
for(i in 1:5) {
  print(i)
}
```

----------------
# Excercise 3C #
----------------

## Paired-sample randomization test

Update this example from class using the following steps

1. Read in `ButterflyFishDiffs.csv`
2. Use the code from lecture to calculate the p-value for a two-sided hypothesis test.
Now run this a couple of times and see how much the p-value varies. 
3. Increase the loop to draw a million samples (1e6) from the null distribution.


```{r}
#update this code

obs.stat <- mean(diff.dat$Difference) 
null.vec <-numeric(1e4)

for(i in 1:length(null.vec)) {
  multiplier <- sample(x=c(-1,1), size=length(diff.dat$Difference), replace=TRUE)
  null.vec[i] <- mean(diff.dat$Difference*multiplier)
}

geom_histogram(aes(x=null.vec), col="white") + xlab("Null hypothesis") +
  geom_vline(aes(xintercept=obs.stat), col='red')
``
```



----------------
# Excercise 3D #
----------------

## Two-sample randomization test

Rerun the hotwings example from lecture (using `hotwings.csv`). Update the code to make a million draws from the null distribution. Does the p-value change from the one I reported?
```{r}
#code here.
```

