---
title: 'Module 5 exercises'
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(ggplot2) #load ggplot
library(ggthemes) #load themes package for ggplot
theme_set(theme_tufte()) # a theme I like.

#set some options that I like
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```

###############
# Exercise 5A #
###############

## Calculating and plotting correlations

Below is the correlation plot of the sleep dataset from lecture:
```{r}
sleep.dat <- read.csv(file="MammalSleep.csv")

library(corrplot)
M <- cor(sleep.dat[,3:10], use="pairwise.complete") #extract numeric responses
corrplot.mixed(M)
```

In lecture we discussed strong correlations between response and predictor. What are the top 3 strongest correlations among the predictor variables in this dataset?

###############
# Exercise 5B #
###############


## Simple linear regression

We are going to repeat the analysis produced in class but instead of using the predictor of `BrainWt` we will transform it and use `log(BrainWt)`. Plug in `log(BrainWt)` to the call in `lm`. What is the R-squared value of this model and how does it compare to the model I presented in class? How about the diagnositic plots, do they look different than the model I presented in class?
```{r}
#code here
```

###############
# Exercise 5C #
###############

## Multiple linear regression

1. Read in the dataset, `monet.csv`. Build a model that includes the effects of SIGNED (whether the painting was signed) and painting SIZE. Based on this, are signed paintings more or less valuable than unsigned paintings?

```{r}
#code here
```

Load the `sjPlot` library and use the function `plot_model` to plot PRICE predictions against SIZE and SIGNED. Does the plot agree with your interpretation of the parameter estimates?

```{r}
#code here
```


###############
# Exercise 5D #
###############

## Run an ANOVA model 

Run an ANOVA on the model you defined above. What does this F-test tell you about the ability of SIGNED to explain variation in the data?
```{r}
#code here
```


