---
title: 'Module 4 exercises'
output: html_notebook
---

```{r}
library(ggplot2) #load ggplot
library(ggthemes) #load themes package for ggplot
theme_set(theme_tufte()) # a theme I like.

#set some options that I like
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```

---------------
# Exercise 4A #
---------------

## Estimating means and standard deviations

Estimate the mean (`mean()`) and standard deviation (`sd()`) for the following data vectors:
```{r}
x <- 0:10
y <- rep(5, 10)
z <- sample(0:10, size=10, replace=TRUE) 

mean(x)
sd(x)

mean(y)
sd(y)

mean(z)
sd(z)

```

Rerun the analysis a couple of times do the estimates from `z` change or stay the same? Does this make sense?

### The mean and standard deviation of `z` changes a bit each time it is run due to the random nature of the sample.

---------------
# Exercise 4B #
---------------

## Calculating z-scores, and p-values 

The example in class on a sample of 80 baby weights in the United States found a mean of $\bar{x}=3370$ with variance of  $\sigma=573$. Calculate the 95% confidence interval for this estimate.

```{r}
#code here
se <- 573/sqrt(80)
xbar <- 3370

xbar + 1.96*c(-1,1)*se
```

In this example the population mean was $\mu=3339$.  Repeat the z-test conducted in class but let the sample size be 160 instead of 80. Does this change the p-value? What if the sample size had been 8000 instead of 80? Would your conclusions about whether this sampled population is the same as the general population change?

```{r}
#code here
z <- (xbar - 3339)/(573/sqrt(160))
2*pnorm(q=z, lower.tail=F)

z <- (xbar - 3339)/(573/sqrt(8000))
2*pnorm(q=z, lower.tail=F)
```


---------------
# Exercise 4C #
---------------

## Calculating Wald confidence intervals 

Update the 95% confidence interval you calculated above to account for uncertainty in $\sigma$ with the t-distribution.

```{r}
#code here
se <- 573/sqrt(80)
xbar <- 3370

xbar + qt(p=0.975, df=80-1)*se
xbar + qt(p=0.025, df=80-1)*se

```
### This interval is slightly wider than the z-interval, but not much.

---------------
# Exercise 4D #
---------------

## Two-sample t-test

Repeat the in-class example by reading in the csv file `BeerGlassShape.csv` and conduding a t-test using the `response ~ predictor` formula.

```{r}
#code here
glass.dat <- read.csv(file="../Data/BeerGlassShape.csv")
t.test(drinkingMinutes ~ glassShape, data=glass.dat)
```

