---
title: 'Module 7 group work'
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(ggplot2) #load ggplot
library(ggthemes) #load themes package for ggplot
theme_set(theme_tufte()) # a theme I like.

#set some options that I like
theme_update(plot.title = element_text(hjust = 0.5), 
             axis.line.x = element_line(color="black", size = 1),
             axis.line.y = element_line(color="black", size = 1),
             text=element_text(size=20),
             axis.text=element_text(size=15)) #center all titles and and axis lines
```


Dodd and Silvertown [1999](https://doi.org/10.1086/303238) explored the relationships between size , age, competition, on the number of cones produced by fir (*Abies balsamea*). Interestingly, some of these populations experience wave-like die offs that lead to bands of dead trees in the forest (see image [here](https://www.conifers.org/pi/ab/balsamea13.jpg)). These waves are thought to be driven by chronic wind-stress. Today we will explore how fir fecundity is related to tree size, and population that have and not had wave die-offs.

- `DBH`: diameter in m at breast height (1.4 m)
- `fecundity`: number of cone rachises produced per year
- `pop`: whether the tree comes from a population that experiences non-wave or wave mortality patterns

1. Read in `Firs.csv` and plot the dataset, illustrating the effect of both `DBH` and population type (`pop`) on fecundity. What kind of distribution should we consider using to model fecundity? Based on this figure do you think the relationship between fecundity and DBH differ between populations?


```{r}
#code here
fir.dat <- read.csv(file="../Data/Firs.csv")
head(fir.dat)

ggplot(data=fir.dat, aes(x=DBH, y=fecundity, col=pop)) + 
  geom_point()

```

2. Now fit a Poisson regression to `fecundity` using `pop` and `DBH` as predictor variables. Report model estimates and confidence intervals, and plot the model fits using the function `plot_model` in the `sjPlot` library.

```{r}
#code here
firs.pois <- glm(fecundity ~ DBH + pop, data=fir.dat, family=poisson)
confint(firs.pois)
summary(firs.pois)
plot_model(firs.pois, type="pred", terms=c("DBH", "pop"), show.data=TRUE)
```

3. Repeat this analysis but use a negative binomial model. Use AICc to determine which fit is better. Do the parameter estimates or confidence intervals differ much between the results of the Poisson and negative binomial distributions? Plot the model predictions from your best model using the `sjPlot` library.

```{r}
#code here
library(MASS)
library(MuMIn)
firs.nb <- glm.nb(fecundity ~ DBH + pop, data=fir.dat)
summary(firs.nb)
confint(firs.nb)
model.sel(list(firs.nb, firs.pois))

plot_model(firs.nb, type="pred", terms=c("DBH", "pop"), show.data=TRUE)
```

### The negative binomial model is much better than the Poisson, as measured by AICc. The predicted means of the two models look very similar, however the error around the predictions is much dwider in the negative binomial model. The coefficeint  for the popwave predictor changes nad the parameter confidence interval widens by an order of magnitude. Thus, the Poisson regression led us to be overconfident about our conclusions that there were differences between these populations. 


----------------------


Let's look at the silversword data from the lecture in more detail.

1. Load the dataset `Silversword.csv`.
2. Build a survival model using the predictors `Pop_size`, `Precip`, `Elev`, and `veg_cover`. 
3. Increases in which predictors are associated with increases in survival? How about decreases in survival? Is there evidence of density dependence in the survival rate?

```{r}
#code here
silver.dat <- read.csv(file="../Data/Silversword.csv")
ggplot(data=silver.dat) +
  geom_point(mapping=aes(x=Elev, y=Survive/Pop_size)) + xlab("Elevation") + ylab("Survival probability")

full.mod <- glm(cbind(Survive, Die) ~ Elev + Precip + veg_cover + Pop_size, family=binomial, data=silver.dat)
library(car)
vif(full.mod)
summary(full.mod)

```

### We see that most of these effects are statistically significant. No obvious indications of problems such as collinearity or overdispersion. 


4 Use the `VGAM` package to fit the model you used above to fit the same mean function as above but use a betabinomial model. 

5. Compare the AIC values between the two model and compare their parameter estimates (you can use the command `coef(mod.name)` ). Are there systematic differences in the estimates? 
*Note that `MuMIn` does not allow for `VGAM` but you can do `AIC(model.name)` to get the value.*

```{r}
#code here
library(VGAM) #contains beta-binomial regression
model.bb <- vglm(cbind(Survive, Die) ~ Elev + Precip + veg_cover + Pop_size, data=silver.dat, family=betabinomial)

coef(full.mod)
summary(model.bb)

AIC(model.bb)
AIC(full.mod)
```
### The beta-binomial is a better fitting model. There are some slight differences in the estimates and precip has gone from having a negative and significant impact on survival to becoming non-signficant. 



