---
title: "Module 1 group work"
output: html_notebook
---

Vectors can be made using the combine function c(). We can then do operations on the vector, instead of each number individually. Multiply the vector below by 10. Then calculate the square of x, $x^2$.

```{r}
x <- c(1, 3, 5, 7, 9, 11)

#code here
x*10
sqrt(x)
x^2
```

Below is a vector of density estimates of the Hawaiʻi ʻelepaio at 12 sites (data from: https://dspace.lib.hawaii.edu/bitstream/10790/2692/1/HCSU012CampetalPopulationtrendsofnativeHawaiianforestbirds1976-2008FINAL.pdf) and the these are the areas surveyed (in km^2) at each site. Calculate the bird density at each site.

```{r}
BirdCounts <- c(15, 44, 96, 790, 0, 23, 158, 9, 118, 38, 59, 16)
SurveyArea <- c(0.40, 0.58, 0.57, 3.45, 0.52, 0.45, 1.12, 0.58, 0.44,0.63 ,0.83, 1.27)

#code here
bird.dens <- BirdCounts/SurveyArea
```

Now calculate the mean and standard deviation in bird density:
```{r}
#code here
mean(bird.dens)
sd(bird.dens)
```

###########

Read in the dataset `ButterflyFish.csv`. 
```{r}
#code here
butt.dat <- read.csv(file="../Data/ButterflyFish.csv")
str(butt.dat)
```

How many variables are in the dataset (columns) and how many observations (rows)?

### 32 observations of 2 variables.

Calculate the mean of the `Length` variable. Try calculating the mean of the `Sex` variable as well. What type of variable is `Sex`?
```{r}
#code here
mean(butt.dat$Length)
mean(butt.dat$Sex)
```
### Calculating the mean of sex doesn't work because it's a factor.

Use the `subset` function to extract the mean and standard deviation of the lengths of male fish. How many observations of male fish are there?

```{r}
#code here
summary(subset(butt.dat, Sex=="Male"))
sd(subset(butt.dat, Sex=="Male")$Length)
```
### There are 16 observations of male fish and the mean is 95.3 and the standard deviation is 7.04.

Use the `subset` function to extract the mean and standard deviation of the lengths of female fish. How many observations of female fish are there?

```{r}
#code here
summary(subset(butt.dat, Sex=="Female"))
sd(subset(butt.dat, Sex=="Female")$Length)
```
### There are 16 observations of female fish and the mean is 93.7 and the standard deviation is 7.52.

